---
title: "sabr - A package to query and calculate MLB offensive performance statistics"
author: "Michael Lee"
date: "April 10, 2016"
output:
  md_document:
    variant: markdown_github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
#sabr

The sabr package connects the Lahman SQL Database with functions used in calculating offensive performance in the MLB, namely wOBA. Created by Tom Tango and popularized in [The Book: Playing the Percentages in Baseball](http://www.insidethebook.com/woba.shtml), wOBA uses yearly MLB performance to weigh offensive measures relative to league value. This package offers the following capabilities:

- Retrieves the most up-to-date wOBA measures from 1871-2015
- Queries yearly and career offensive performance from the Lahman SQL database
- Calculates a player's seasonal wOBA performance

## Installation:

* the latest development version from github with

    ```R
    if (packageVersion("devtools") < 1.6) {
      install.packages("devtools")
    }
    devtools::install_github("mdlee12/sabr")
    ```

## Motivation:

wOBA is one of the best catch-all offensive performance statistics, but it's formula is a bit convoluted since it requires 1) ten distinct rate stats and 2) weighted values based on a season's run environment. 

For example, the wOBA formula for the 2013 season was:

wOBA = (0.690×uBB + 0.722×HBP + 0.888×1B + 1.271×2B + 1.616×3B +
2.101×HR) / (AB + BB – IBB + SF + HBP)

Check out [Fangraphs](http://www.fangraphs.com/library/offense/woba/)' library for additional details.

## New Functions with 0.2:

* wRAA_season : Returns a MLB player's weighted Runs Above Average by yearID
```{r}
library(sabr)
wRAA_season("braunry02","2015")
```

* wRC_season : Returns a MLB player's weighted Runs Created by yearID
```{r}
wRC_season("molitpa01","1991")
```

* wRCp_season : Returns a MLB player's weighted Runs Created Plus by yearID
```{r}
wRCp_season("yountro01","1989")
```
*Full calculation details are in the [wOBA and wRC+ calculation.Rmd file](https://github.com/mdlee12/sabr/vignettes/wOBA and wRC+ calculation.Rmd)*
 + wRCp_season makes a couple assumptions:
  1. Park Factors are pulled from the Lahman Database - these are different from other sources
  2. A player's primary position is used to remove pitchers from the League Runs/Plate Appearance calculation

******

## Key Functions from 0.1:

* battingseason : Offensive seasonal performance by playerID and yearID
```{r}
battingseason("parrage01","2014")
```

* battingcareer : Offensive career performance by playerID by organized by season
```{r}
battingcareer("fieldpr01")
```

* woba_season : Returns a MLB player's weighted On-Base Average by yearID
```{r}
woba_season("fieldpr01","2014")
```


* lahman_search : Returns a player's Lahman playerID using either:
 + nameFirst (Player First Name)
 + nameLast (Player Last Name)
 + nameGiven (Player Given Name)
 + retroID ([Retrosheet](http://www.retrosheet.org/retroID.htm) ID)
 + bbrefID ([Baseball Reference](http://www.baseball-reference.com/players/) ID)
```{r}
lahman_search(nameFirst = "Robin")
lahman_search(nameLast = "Fielder")
lahman_search(nameGiven = "Russell Oles")
lahman_search(retroID = "molip001")
lahman_search(bbrefID = "yostne01")
```

<<<<<<< HEAD
Have a question, issue or suggestion? Create a pull request, file an issue, or feel free to contact me on [Twitter](https://twitter.com/mikeleeco) or [my website](http://mikelee.co/)!
=======
Have a question, issue or suggestion? Create a pull request or feel free to contact me on [Twitter](https://twitter.com/mikeleeco)!
>>>>>>> origin/master
